A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Edits to the two Content Recommender sections

\subsection*{\em ContentBasedRecommender}
The most straightforward approach to inter-retailer recommendations is a
content-based approach. Retail products typically have a description that can be
used to construct a textual content-based similarity function. One can then
consider all pairwise product combinations across retailers, but within a common
category, and choose recommendations between items with sufficiently high
similarity.

Toward this end, we borrow the concept of {\em Term Frequency--Inverse Document
Frequency} (tfidf) from the information retrieval community. In using tf-idf, we
are considering each item description to be a ``document''.  Thus we define the
tfidf of a term $t$ within an item description $d$ that's part of a combined
category catalogue $C$ (i.e. the catalogue consisting of all items in both
retailers' catalogues for a particular common category) as
\begin{align}
\mathrm{tfidf}(t,d,C) = f(t,d) \times
\mathrm{log}\frac{\left|C\right|}{\left|\left\{d \in C : t \in
d\right\}\right|},
\end{align}
where $f(t,d)$ is the number of occurances of term $t$ in item description $d$,
and $\left|\left\{d \in C : t \in d\right\}\right|$ is the number of items in
$C$ with descriptions containing at least one instance of term $t$.  We can now
represent each item as a sparse vector of tfidf values, which has a nonzero
element for each term in its description.

A key feature of tfidf that is useful here is how terms that appear in many item
descriptions are discounted via the idf factor. Conversely, rare terms will have
larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
$B$ as
\begin{align}
ContentSimilarity(A,B) = v_A \cdot v_B
\end{align}
where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
respectively.

Using this content-based similarity function, we construct the following
inter-catalogue recommendation algorithm, which we call the $ContentBased
Recommender$. For each product in each catalogue, calculate its
$ContentSimilarity$ with each and every product in the other catalogue. Predict
recommendations between this product and the top two most similar products in
the other catalogue. The reason for choosing the top two most similar products
originates from our knowledge that most items listed on www.macy.com are
associated with four recommended items. Thus, by choosing two edges per item per
partition, we will predict approximately the same number of edges that were lost
during the partition. This relatively straightforward algorithm serves as a
baseline against which to compare the performance of other more sophisticated
algorithms.

The $ContentBasedRecommender$ deserves a few more comments. For starters, the
use of an inner product, which is bounded only by the length of the tfidf
vectors, is not fundamental to this algorithm. A bounded similarity function
such as cosine similarity could have been used instead without qualitatively
changing the results discussed later in this report. This is because the lengths
of the product descriptions have little variation, and therefore, the vector
norms of the tfidf vectors have little variation. Finally, we mention that while
this algorithm is quadratic in the number items, it can be executed relatively
efficiently using matrix multiplication, which trades time for space (i.e.
memory). Furthermore, confining our experiments to individual categories reduces
the impact of the quadratic time complexity. In our case, Macy's entire
catalogue consists of over $60,000$ products, making a quadratic time algorithm
challenging. However, the average category has about two thousand products,
which is significantly more manageable.

\subsection*{\em NeighborhoodRecommender}
The $ContentBasedRecommender$ is limited in that it attempts to capture
similarity between products by looking exclusively at the terms in their
descriptions. One issue here is that brands tend to use different vocabularies
to describe their products. One may refer to a color as ``off-white'' while
another calls the same color ``ivory.'' To the $ContentBasedRecommender$ these
terms are as different as ``black'' and ``white.'' Another issue is that the
best recommendation for a product may not even be the most similar product. For
instance, a striped black-and-white button-up shirt could be an excellent
recommendation associated with a blue checkered button-up, even though the two
products are clearly different and would presumably use signficantly different
terms their descriptions.

@@

Assuming that the edges of the recommendation graphs are in fact good
recommendations, this approach can help both of the shortcomings of
$ContentBasedRecommender$ previously mentioned. Since a neighborhood includes
several items, each with descriptions that potentially use different
vocabularies, there is a higher likelihood that the neighborhood contains
multiple synonyms for the underlying concept that captures their similarity. At
the same time, this decreases the likelihood of false negatives due to different
vocabularies. Presumably, a retailer's recommendation graph include information
from user behavior that is not expressed in product descriptions. Re-examining
the button-up shirts example, the neighborhood of a striped shirt may include
checkered shirts and vice versa.  In this case, the two neighborhoods might
share the words ``striped'' and ``checkered'' and thus the two previously
dissimilar shirts would now be considered more similar.

@@

To map concepts from one catalogue to another, we must revisit
content-similarity. In this case, we represent each concept (rather than each
product) as a vector of words. To represent a concept in terms of the
vocabulary, we calculate the value of each word by summing the
tf-idf of that word for each product $P$, weighted by the projection of $P$ on
our concept in concept-space. That is:

\begin{align}
C_w = \sum\limits_{P} \text{tfidf}(w)\times P_C
\end{align}
where $C_w$ is the value of concept $C$'s representation in dimension $w$, the
sum is over all products $P$ in the catalgoue, tfidf($w$) is the tf-idf score
for word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
concept-space.

>>>>

\subsection*{\em ContentBasedRecommender}
The most straightforward approach to inter-retailer recommendations is a
content-based approach. Retail products typically have a description that can be
used to construct a textual content-based similarity function. One can then
consider all pairwise product combinations across the two retailers, but within
a common category, and choose recommendations between items with sufficiently
high similarity.

Toward this end, we borrow the concept of {\em Term Frequency--Inverse Document
Frequency} (tfidf) from the information retrieval community. In using tfidf, we
are considering each item description to be a ``document''.  Thus we define the
tfidf of a term $t$ within an item description $d$ that's part of a combined
category catalogue $C$ (i.e. the catalogue consisting of all items in both
retailers' catalogues for a particular common category) as
\begin{align}
\mathrm{tfidf}(t,d,C) = f(t,d) \times
\mathrm{log}\frac{\left|C\right|}{\left|\left\{d \in C : t \in
d\right\}\right|},
\end{align}
where $f(t,d)$ is the number of occurances of term $t$ in item description $d$,
and $\left|\left\{d \in C : t \in d\right\}\right|$ is the number of items in
$C$ with descriptions containing at least one instance of term $t$.  We can now
represent each item as a sparse vector of tfidf values, which has a nonzero
element for each term in its description.

A key feature of tfidf is that terms appearing in many item descriptions are
discounted via the idf factor. Conversely, rare terms will have larger tfidf
values. These features help capture the most relevant words in each description.
We now define the $ContentSimilarity$ between items $A$ and $B$ as
\begin{align}
ContentSimilarity(A,B) = v_A \cdot v_B
\end{align}
where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
respectively.

Using this content-based similarity function, we construct the following
inter-catalogue recommendation algorithm, which we call the $ContentBased
Recommender$. For each product in each catalogue, calculate its
$ContentSimilarity$ with every product in the other catalogue. Predict
recommendations between this product and the two most similar products in
the other catalogue. The reason for choosing the top two most similar products
originates from our knowledge that most items listed on www.macy.com are
associated with four recommended items. Thus, by choosing two edges per item per
partition, we will predict approximately the same number of edges that were lost
during the partition. This relatively straightforward algorithm serves as a
baseline against which to compare the performance of other more sophisticated
algorithms.

The $ContentBasedRecommender$ deserves a few more comments. For starters, the
use of an inner product, which is bounded only by the length of the tfidf
vectors, is not fundamental to this algorithm. A bounded similarity function
such as cosine similarity could have been used instead, without qualitatively
changing the results discussed later in this report. This is because the lengths
of the product descriptions have little variation, and therefore, the vector
norms of the tfidf vectors have little variation. Finally, we mention that while
this algorithm is quadratic in the number of items, it can be executed relatively
efficiently using matrix multiplication, which trades time for space (i.e.
memory). Furthermore, confining our experiments to individual categories reduces
the impact of the quadratic time complexity. In our case, Macy's entire
catalogue consists of over $60,000$ products, making a quadratic time algorithm
challenging. However, the average category has about two thousand products,
which is significantly more manageable.

\subsection*{\em NeighborhoodRecommender}
The $ContentBasedRecommender$ is limited in that it attempts to capture
similarity between products by looking exclusively at the terms in their
descriptions. One issue here is that brands tend to use different vocabularies
to describe their products. One may refer to a color as ``off-white'' while
another calls the same color ``ivory.'' To the $ContentBasedRecommender$ these
terms are as different as ``black'' and ``white.'' Another issue is that the
best recommendation for a product may not even be the most similar product. For
instance, a striped black-and-white button-up shirt could be an excellent
recommendation associated with a blue checkered button-up, even though the two
products are different and would presumably use signficantly different terms in
their descriptions.

@@

Assuming that the edges of the recommendation graphs are in fact good
recommendations, this approach can help both of the shortcomings of
$ContentBasedRecommender$ previously mentioned. Since a neighborhood includes
several items, each with descriptions that potentially use different
vocabularies, there is a higher likelihood that the neighborhood contains
multiple synonyms for the underlying concept that captures their similarity.
This therefore decreases the likelihood of false negatives due to different
vocabularies between a pair of items. Presumably, a retailer's recommendation
graph includes information from user behavior that is not expressed in product
descriptions. Re-examining the button-up shirts example, the neighborhood of a
striped shirt may include checkered shirts and vice versa.  In this case, the
two neighborhoods might share the words ``striped'' and ``checkered'' and thus
the two previously dissimilar shirts would now be considered more similar.

@@

To map concepts from one catalogue to another, we must revisit
content-similarity. In this case, we represent each concept (rather than each
product) as a vector of words. To represent a concept in terms of the
vocabulary, we calculate the value of each word by summing the
tfidf of that word for each product $P$, weighted by the projection of $P$ on
our concept in concept-space. That is:

\begin{align}
C_w = \sum\limits_{P} \text{tfidf}(w)\times P_C
\end{align}
where $C_w$ is the value of concept $C$'s representation in dimension $w$, the
sum is over all products $P$ in the catalgoue, tfidf($w$) is the tfidf score
for word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
concept-space.
